---
output:
  pdf_document: default
  html_document: default
---


RepData_PeerAssessment1
========================================

Loading the data
```{r echo=FALSE}
setwd("~/Downloads/")
```

```{r}
data <- read.csv2("activity.csv", header= TRUE, sep = ",")
summary(data)
head(data)

```
```{r}
data2 = as.data.frame(data[complete.cases(data), ])
summary(data)
```


# Histogram of the total number of steps taken each day

```{r}
steps <- tapply(data2$steps, data2$date, sum)

hist(steps,main = "Number of steps per Day", xlab = "Total Steps per Day", ylab = "Days")

```

#Mean and median number of steps taken each day
```{r}
meansteps = mean(steps, na.rm = TRUE)
meansteps
```

```{r}
mediansteps = median(steps, na.rm = TRUE)
mediansteps
```

#Time series plot of the average number of steps taken
```{r}
avgerage_number = aggregate(data2$steps, by = list(data2$interval), FUN = mean)
summary(avgerage_number)
```

```{r}
plot(avgerage_number[, 1], avgerage_number[, 2], type = "l", 
     xlab = "5 min Intervals in a day", 
     ylab = "Average Number of Steps", 
     main = "The Average Daily Activity Pattern")
```


#The 5-minute interval that, on average, contains the maximum number of steps

```{r}
maxsteps = avgerage_number[which.max(avgerage_number[, 2]), 1]
maxsteps
```

#Code to describe and show a strategy for imputing missing data

```{r}
sum(!complete.cases(data))
```

```{r}
data3 = data
len1 = nrow(data3)
len2 = nrow(avgerage_number)
for (i in 1:len1) {
  if (is.na(data3$steps[i])) {
    for (j in 1:len2) {
      if (data3$interval[i] == avgerage_number[j, 1]) {
        data3$steps[i] = avgerage_number[j, 2]
      }
    } 
  }    
}
summary(data3)
```

#Histogram of the total number of steps taken each day after missing values are imputed
```{r}
totalsteps2 = tapply(data3$steps, data3$date, sum)
hist(totalsteps2, main = "Number of steps per Day", xlab = "Total Steps per Day", ylab = "Days" )
```

#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
data2$weekday = TRUE
weekday = weekdays(as.POSIXct(data2$date, format = "%Y-%m-%d" ))
for (i in 1:length(weekday)) {
  if (weekday[i] == "Saturday" | weekday[i] == "Sunday") {
    data2$weekday[i] = FALSE
  }
}
data_weekday = data2[which(data2$weekday == TRUE), ]
data_weekend = data2[which(data2$weekday == FALSE), ]

average_weekday = aggregate(data_weekday$steps, 
                               by = list(data_weekday$interval), 
                               FUN = mean)
names(average_weekday) = c("interval", "steps")
average_weekday$dayTag = "weekday"

average_weekend = aggregate(data_weekend$steps, 
                              by = list(data_weekend$interval), 
                              FUN = mean)
names(average_weekend)= c("interval", "steps")
average_weekend$dayTag = "weekend"

average_pattern = rbind(average_weekday, average_weekend)
```

```{r echo=FALSE}
library(lattice)
```


```{r}
xyplot(steps ~ interval | dayTag, data = average_pattern, 
       type = "l", layout = c(1, 2))
```

